---
title: "Create preliminary figures for BES poster"
author: "Eleanor Jackson"
date: '`r format(Sys.time(), "%d %B, %Y")`'
always_allow_html: true
output: 
  github_document:
    keep_html: true
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.path = "figures/2022-11-22_create-poster-figures/")
```

```{r packages}
library("tidyverse"); theme_set(theme_bw())
library("here")
library("sf")
library("ggmap")
library("patchwork")
library("ggtext")
```

## Map of conspecific density across BCI

```{r read-gpx}
plotKML::readGPX(here::here("data", "maps",
                            "jacc-map-garzonlopez2012", "jac1co_map.gpx")) %>%
  map_df( ~ .) %>%
  select(lon, lat) %>%
  mutate(id = paste0("CGL", 1:n())) -> carol_jacc

read_sf(here::here("data", "maps", "BCI_Plot_50ha", 
                                "BCI_Plot_50ha.shp")) %>%
  st_transform(plot_50ha, crs = st_crs(4326)) -> plot_50ha

```

```{r jacc-map, message=FALSE}
bbox <- make_bbox(c(min(carol_jacc$lon) - 0.001, max(carol_jacc$lon) + 0.005), 
                  c(min(carol_jacc$lat) - 0.005, max(carol_jacc$lat) + 0.001))

bci_basemap <- ggmap::get_map(bbox, source = "stamen", 
                       force = TRUE, maptype = "toner-lite")

```

```{r density-map}
bci_basemap %>% 
  ggmap() +
  stat_density2d_filled(data = carol_jacc,
                 aes(lon, lat, fill = ..level..), 
                 alpha = 0.7, show.legend = FALSE, 
                 contour_var = "count", bins = 20) +
  geom_sf(data = plot_50ha, inherit.aes = FALSE, fill = NA, colour = "red") +
  theme_void()

```

### Map of conspecific density in the 50ha plot

```{r}
readRDS("/Users/eleanorjackson/OneDrive - University of Reading/spatial-premature-fruit-drop/data/clean/tree_data.rds") %>%
  filter(sp4 == "jacc" & year == 2017) %>%
  rename(lon = x, lat = y) -> ha50_trees

ggplot() +
  stat_density2d_filled(data = ha50_trees,
                 aes(lon, lat, fill = ..level..), alpha = 0.7, 
                 geom = "polygon", show.legend = FALSE) +
  geom_point(data = ha50_trees, aes(lon, lat))+
  ylim(0, 500) +
  xlim(0, 1000) +
  theme_void()

```

## Connectivity histograms

### whole BCI connectivity histogram

```{r read-gpx}
plotKML::readGPX(here::here("data", "raw", "Final_JACC_sites.gpx")) %>%
  map_df( ~ .) %>%
  rename(id = name) %>%
  arrange(id) %>%
  select(id, lon, lat) %>%
  filter(!grepl("NS", id)) %>% 
  add_row(id = "EJJACC155", lon = -79.837979, lat = 9.171755) %>%
  add_row(id = "EJJACC156", lon = -79.828260, lat = 9.160488) %>%
  add_row(id = "EJJACC157", lon = -79.827505, lat = 9.161556) %>%
  add_row(id = "EJJACC119", lon = -79.851366, lat = 9.152496) -> focal_jacc

```

Remove any trees in Carol's data that could be duplicates of our focal trees

```{r}
geosphere::distm(cbind(pull(focal_jacc, lon), pull(focal_jacc, lat)), 
                 cbind(pull(carol_jacc, lon), pull(carol_jacc, lat)), 
                 fun = distGeo) -> dist_matrix

rownames(dist_matrix) <- focal_jacc$id
colnames(dist_matrix) <- carol_jacc$id

as.data.frame.table(dist_matrix, responseName = "dist") %>% 
  filter(dist <= 30) %>%
  group_by(Var1) %>%
  slice(which.min(dist)) %>% 
  ungroup() %>%
  pull(Var2) %>%
  unique() -> duplicate_trees

rbind(carol_jacc, focal_jacc) %>% 
  filter(!id %in% duplicate_trees) -> all_jacc

```

Calculate distances between all trees

```{r}

calculate_dist <- function (data) {

	data %>%
    select(lat, lon) -> plot_matrix

	rdist::pdist(plot_matrix[,c("lat", "lon")], 
	             metric = "euclidean") -> dists
	
	as.data.frame(dists) -> dists_df

	unlist(data$id) -> colnames(dists_df) 

	cbind(data, dists_df)
	
}

distance_df <- calculate_dist(all_jacc)

```

Calculate connectivity of all trees

```{r}
all_jacc %>% 
  distinct(id) %>%
  pull(id) -> tree_id_list

calculate_connectivity <- function (data, tree_id) {
  data %>%
    group_by(id) %>%
    mutate(x = exp(- eval(parse(text = tree_id)) ) ) %>%
    ungroup() %>%
    summarise(id = paste(tree_id),
              connectivity = sum(x), .groups = "drop")
}

connectivity_dfs <- lapply(tree_id_list,
                           calculate_connectivity, data = distance_df)

connectivity_dfs %>%
  lapply(drop_na, connectivity) %>%
  dplyr::bind_rows() -> all_connectivity_dfs

```

Plot histogram

```{r histogram}
all_connectivity_dfs %>%
  mutate(focal = grepl("EJJACC", id)) %>%
  filter(focal == TRUE) %>%
  select(id, connectivity) %>%
  mutate(study = "whole island") -> bci_focal_connect

bci_focal_connect %>%
  ggplot() +
  geom_histogram(aes(x = connectivity))

```

### 50 ha plot connectivity histogram

```{r}
# filter for trees in carol's maps that are within the 50ha
read_sf(here::here("data", "maps",
                            "jacc-map-garzonlopez2012", "JAC1COpointSept.shp")) %>%
  st_transform(crs = st_crs(4326)) %>%
  sf::st_intersection(plot_50ha) %>%
  ggplot(aes(x = X, y = Y))+
  stat_density2d_filled(aes(fill = ..level..), alpha = 0.7, 
                 geom = "polygon", show.legend = FALSE) +
  geom_point() +
  theme_void()

```

Carol's data looks similar to the 50ha plot data above - a good sign.

```{r}
read_sf(here::here("data", "maps",
                            "jacc-map-garzonlopez2012", "JAC1COpointSept.shp")) %>%
  mutate(id = paste0("CGL", 1:n())) %>%
  st_transform(crs = st_crs(4326)) %>%
  sf::st_intersection(plot_50ha) %>%
  left_join(all_connectivity_dfs, by = "id") -> plot_50ha_connect

ggplot(plot_50ha_connect) +
  geom_histogram(aes(x = connectivity))

```

### Comparison histogram

```{r}
plot_50ha_connect %>%
  st_drop_geometry() %>%
  select(id, connectivity) %>%
  mutate(study = "50ha plot") %>%
  rbind(bci_focal_connect) %>%
  ggplot(aes(x = connectivity, fill = study)) +
  geom_histogram(position = "identity",
                 alpha = 0.8,
                 bins = 50) +
  scale_color_manual(values = c("#999999", "#440154FF")) +
  scale_fill_manual(values = c("#999999",  "#440154FF")) +
  theme_classic(base_size = 10) +
  scale_x_continuous(expand = c(0, 0)) +
  xlab("conspecific density") +
  ylab("") +
  labs(subtitle = "n trees") + 
  scale_y_continuous(breaks = seq(0, 60, 10)) +
  theme(
    line = element_blank(),
    axis.text.x = element_blank(),
    legend.position = "none",
    plot.subtitle = element_text(colour = "#2c365e", hjust = -0.05, vjust = -5),
    axis.text.y = element_text(colour = "#2c365e"),
    axis.title.x = element_text(colour = "#2c365e")
  ) -> p


p +
  annotate(
    geom = "curve",
    x = 1051, y = 40,
    xend = 1054, yend = 20,
    colour = "#999999",
    curvature = .3,
    arrow = arrow(length = unit(2, "mm"))
  ) +
  annotate(
    geom = "text",
    x = 1052, y = 47,
    colour = "#999999",
    label = "trees sampled in the\n 50ha plot (0.5km2)",
    hjust = "right",
    size = 4
  ) +
  
  annotate(
    geom = "curve",
    x = 1048, y = 30,
    xend = 1051, yend = 5,
    colour = "#440154FF",
    curvature = -.2,
    arrow = arrow(length = unit(2, "mm"))
  ) +
  annotate(
    geom = "text",
    x = 1047.7, y = 30,
    colour = "#440154FF",
    label = "trees sampled in our study\n across the 15.6km2 island",
    hjust = "right",
    size = 4
  ) +
  theme(plot.margin = margin(1, 1, 1, 1, "mm")) +
  coord_cartesian(ylim = c(0, 65), clip = 'off') -> p2

ggsave(here::here("01_histogram_new2.png"), plot = p2,
    device = "png", dpi = 600, width = 120, height = 80, units = "mm")
  
```

## proportion pods predated ~ connectivity

```{r}
read.csv(here::here("data", "raw", "jacaranda_pods.csv")) -> pod_data

pod_data %>%
  count(tree, pod_predated) %>%
  spread(pod_predated, n) %>%
  rename(n_predated = `TRUE`) %>%
  mutate(`FALSE` = replace_na(`FALSE`, 0),
         n_predated = replace_na(n_predated, 0)) %>%
  summarise(
    tree = tree,
    n_predated = n_predated,
    n_capsules = n_predated + `FALSE`,
    proportion_predated = n_predated / (n_predated + `FALSE`)
  ) -> porportion_pods

bci_focal_connect %>%
  mutate(id = gsub("EJ", "", id) ) %>%
  mutate(id = gsub("^(JACC)", "\\1_\\2", id)) %>% 
  right_join(porportion_pods, by = c("id" = "tree")) -> pods_connect

pod_data %>% 
  select(tree, dbh_mm, pod_size_mm, morph) %>% 
  distinct() %>%
  right_join(pods_connect, by = c("tree" = "id")) -> pods_connect_dbh

glm(pods_connect_dbh, formula = cbind(n_predated, (n_capsules - n_predated)) ~ connectivity,
    family = "binomial") -> m1
summary(m1)

pods_connect_dbh %>%
  ggplot(aes(
    x = connectivity,
    y = n_predated / n_capsules,
    succ = n_predated,
    fail = (n_capsules - n_predated)
  )) +
  geom_point(
    colour = "#440154FF",
    alpha = 0.7,
    shape = 16) +
  geom_smooth(
    method = "glm", 
    colour = "#2c365e",
    se = TRUE,
    method.args = list(family = "binomial"),
    formula = cbind(succ, fail) ~ x
  ) +
  theme_classic(base_size = 10) +
  xlab("conspecific density") +
  ylab("") +
  scale_y_continuous(breaks = seq(0, 0.6, 0.1)) +
  theme(
    line = element_blank(),
    axis.text.x = element_blank(),
    legend.position = "none",
    axis.text.y = element_text(colour = "#2c365e"),
    axis.title.x = element_text(colour = "#2c365e"),
    plot.subtitle = element_text(colour = "#2c365e", hjust = -0.07, vjust = -6),
    plot.caption = element_markdown(colour = "#2c365e", size = 5)
  ) +
  theme(plot.margin = margin(1, 1, 1, 1, "mm")) +
  coord_cartesian(ylim = c(0, 0.65), clip = 'off') +
  labs(caption = paste("binomial glm: β = ", 
                       parse(text = round(pluck(m1$coefficients, 2), digits = 2)),
                       " [", parse(text = round(pluck(confint(m1, "connectivity"), 1), digits = 2)), ", ", 
                       parse(text = round(pluck(confint(m1, "connectivity"), 2), digits = 2)), "],",
                       " _p_ = < 0.001", sep = ""),
       subtitle = "propotion of \npredated fruits") -> p3

ggsave(here::here("01_glm.png"), plot = p3,
    device = "png", dpi = 600, width = 120, height = 80, units = "mm")

```

## n pods ~ connectivity

```{r}
glm(formula = n_capsules ~ connectivity, data = pods_connect_dbh) -> m2
summary(m2)

pods_connect_dbh %>%
  ggplot(aes(
    x = connectivity,
    y = n_capsules)) +
  geom_point(
    colour = "#440154FF",
    alpha = 0.7,
    shape = 16) +
  geom_smooth(
    method = "glm", 
    colour = "#2c365e",
    se = TRUE
  ) +
  theme_classic(base_size = 10) +
  xlab("conspecific density") +
  ylab("") +
  theme(
    line = element_blank(),
    axis.text.x = element_blank(),
    legend.position = "none",
    axis.text.y = element_text(colour = "#2c365e"),
    axis.title.x = element_text(colour = "#2c365e"),
    plot.subtitle = element_text(colour = "#2c365e", hjust = -0.07, vjust = -3),
    plot.caption = element_markdown(colour = "#2c365e", size = 5)
  ) +
  theme(plot.margin = margin(1, 1, 1, 1, "mm")) +
  coord_cartesian(ylim = c(0, 500), clip = 'off') +
  labs(caption = paste("gaussian glm: β = ", 
                       parse(text = round(pluck(m2$coefficients, 2), digits = 2)),
                       " [", parse(text = round(pluck(confint(m2, "connectivity"), 1), digits = 2)), ", ", 
                       parse(text = round(pluck(confint(m2, "connectivity"), 2), digits = 2)), "],",
                       " _p_ = 0.136", sep = ""),
       subtitle = "n fruits") -> p4

ggsave(here::here("02_glm.png"), plot = p4,
    device = "png", dpi = 600, width = 120, height = 80, units = "mm")

```

## porportion pods aborted ~ connectivity

```{r}
pod_data %>%
  mutate(pod_size_mm = as.numeric(pod_size_mm)) %>%
  filter(pod_size_mm < 200) %>% 
  mutate(pod_aborted = case_when(pod_size_mm < 40 &
                               morph == "symmetrical_locules" ~ TRUE,
                             is.na(pod_size_mm) ~ NA,
                             TRUE ~ FALSE)) %>% 
  count(tree, pod_aborted) %>%
  spread(pod_aborted, n) %>% 
  rename(n_aborted = `TRUE`) %>%
  mutate(`FALSE` = replace_na(`FALSE`, 0),
         n_predated = replace_na(n_aborted, 0)) %>%
  summarise(
    tree = tree,
    n_aborted = n_aborted,
    n_capsules = n_aborted + `FALSE`,
    proportion_aborted = n_aborted / (n_aborted + `FALSE`)
  ) -> porportion_pods_aborted

porportion_pods_aborted %>%
  left_join(pods_connect, by = c("tree" = "id")) -> clean_data_aborted

glm(clean_data_aborted, formula = cbind(n_aborted, (n_capsules.x - n_aborted)) ~ connectivity,
    family = "binomial") -> m3
summary(m3)

clean_data_aborted %>%
  ggplot(aes(
    x = connectivity,
    y = n_aborted / n_capsules.x,
    succ = n_aborted,
    fail = (n_capsules.x - n_aborted)
  )) +
  geom_point(
    colour = "#440154FF",
    alpha = 0.7,
    shape = 16) +
  geom_smooth(
    method = "glm", 
    colour = "#2c365e",
    se = TRUE,
    method.args = list(family = "binomial"),
    formula = cbind(succ, fail) ~ x
  ) +
  theme_classic(base_size = 10) +
  xlab("conspecific density") +
  ylab("") +
  scale_y_continuous(breaks = seq(0, 0.7, 0.1)) +
  theme(
    line = element_blank(),
    axis.text.x = element_blank(),
    legend.position = "none",
    axis.text.y = element_text(colour = "#2c365e"),
    axis.title.x = element_text(colour = "#2c365e"),
    plot.subtitle = element_text(colour = "#2c365e", hjust = -0.07, vjust = -6),
    plot.caption = element_markdown(colour = "#2c365e", size = 5)
  ) +
  theme(plot.margin = margin(1, 1, 1, 1, "mm")) +
  coord_cartesian(ylim = c(0, 0.75), clip = 'off') +
  labs(caption = paste("binomial glm: β = ", 
                       parse(text = round(pluck(m3$coefficients, 2), digits = 2)),
                       " [", parse(text = round(pluck(confint(m3, "connectivity"), 1), digits = 2)), ", ", 
                       parse(text = round(pluck(confint(m3, "connectivity"), 2), digits = 2)), "],",
                       " _p_ = < 0.001", sep = ""),
       subtitle = "propotion of \naborted fruits") -> p5

ggsave(here::here("03_glm.png"), plot = p5,
    device = "png", dpi = 600, width = 120, height = 80, units = "mm")

```

## pod size ~ connectivity

```{r}
pods_connect_dbh %>%
  mutate(pod_size_mm = as.numeric(pod_size_mm)) %>%
  filter(morph == "symmetrical_locules") %>% 
  glm(formula = pod_size_mm ~ connectivity,
    family = "gaussian") -> m4
summary(m4)

pods_connect_dbh %>%
  mutate(pod_size_mm = as.numeric(pod_size_mm)) %>%
  filter(morph == "symmetrical_locules") %>% 
  ggplot(aes(
    x = connectivity,
    y = pod_size_mm
  )) +
  geom_jitter(
    colour = "#440154FF",
    alpha = 0.5,
    shape = 16) +
  geom_smooth(
    method = "glm", 
    colour = "#2c365e",
    se = TRUE
  ) +
  theme_classic(base_size = 10) +
  xlab("conspecific density") +
  ylab("") +
  theme(
    line = element_blank(),
    axis.text.x = element_blank(),
    legend.position = "none",
    axis.text.y = element_text(colour = "#2c365e"),
    axis.title.x = element_text(colour = "#2c365e"),
    plot.subtitle = element_text(colour = "#2c365e", hjust = -0.07, vjust = -3),
    plot.caption = element_markdown(colour = "#2c365e", size = 5)
  ) +
  theme(plot.margin = margin(1, 1, 1, 1, "mm")) +
  coord_cartesian(ylim = c(0, 150), clip = 'off') +
  labs(caption = paste("gaussian glm: β = ", 
                       parse(text = round(pluck(m4$coefficients, 2), digits = 2)),
                       " [", parse(text = round(pluck(confint(m4, "connectivity"), 1), digits = 2)), ", ", 
                       parse(text = round(pluck(confint(m4, "connectivity"), 2), digits = 2)), "],",
                       " _p_ = < 0.001", sep = ""),
       subtitle = "fruit size (mm)") -> p6

ggsave(here::here("04_glm.png"), plot = p6,
    device = "png", dpi = 600, width = 120, height = 80, units = "mm")
```


## Map of seed traps

```{r}
readRDS("/Users/eleanorjackson/OneDrive - University of Reading/spatial-premature-fruit-drop/data/clean/trap_data.rds") %>%
  filter(year == "2018") %>% 
  distinct(trap, x, y) -> trap_loc

ggplot(trap_loc, aes(x = x, y = y)) +
  geom_point(
    colour = "#440154FF",
    alpha = 0.7,
    shape = 16) +
  theme_classic(base_size = 10) +
  labs(x = "1000m", y = "500m") +
  theme(
    line = element_blank(),
    axis.title.y = element_text(colour = "#2c365e", angle = 0),
    axis.title.x = element_text(colour = "#2c365e", hjust = 1),
    axis.text.y = element_blank(),
    axis.text.x = element_blank(),
    axis.line.y = element_line(arrow = grid::arrow(length = unit(0.2, "cm"), 
                                                       ends = "last"), 
                               colour = "#2c365e", size = 0.5),
    axis.line.x = element_line(arrow = grid::arrow(length = unit(0.2, "cm"), 
                                                       ends = "last"), 
                               colour = "#2c365e", size = 0.5)
  ) -> p4
  
ggsave(here::here("03_trap_map.png"), plot = p4,
    device = "png", dpi = 600, width = 120, height = 80, units = "mm")

```

