---
title: "Explore conspecific densities for focal trees"
author: "Eleanor Jackson"
date: '`r format(Sys.time(), "%d %B, %Y")`'
always_allow_html: true
output: 
  github_document:
    keep_html: true
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.path = "figures/2022-10-28_explore-conspecific-densities/")
library("emo")
```

We have found ~80 _Jacaranda copaia_ individuals across the island, but have we got a good range of conspecific densities in there? Do we need to find more trees to fill in any gaps? 

We could use Carol's _Jacaranda_ maps to look at natural densities of the population and overlay with the densities of our chosen individuals.

```{r packages}
library("tidyverse"); theme_set(theme_bw())
library("here")
library("sf")
library("ggmap")
```

## Read in gpx files

```{r read-gpx}
plotKML::readGPX(here::here("data", "maps", "focal_jacc.gpx")) %>%
  map_df( ~ .) %>%
  rename(id = name) %>%
  arrange(id) %>%
  select(id, lon, lat) -> focal_jacc

plotKML::readGPX(here::here("data", "maps",
                            "jacc-map-garzonlopez2012", "jac1co_map.gpx")) %>%
  map_df( ~ .) %>%
  select(lon, lat) %>%
  mutate(id = paste0("CGL", 1:n())) -> carol_jacc

rbind(carol_jacc, focal_jacc) -> all_jacc
```

To calculate connectivity you need an estimate of 'patch size' -  I've just realised that we don't have this for Carol's mapped trees and we haven't finished collecting dbh for our focal trees yet. Will have to use some kind of simpler metric. 

Looking in [Moilanen & Nieminen 2002](https://doi.org/10.1890/0012-9658(2002)083[1131:SCMISE]2.0.CO;2), the simplest connectivity measure is isolation (equation 1a), where isolation of the focal tree is equal to the distance from it's nearest neighbour. I don't think this will work for us as we want to look on a bigger scale and take into account all trees across the island. We could calculate Hanski's connectivity without alpha so that the buffer radius = infinity and also remove `A` (patch size).

Hanski's connectivity:

$$
S_i = \sum exp(- \alpha dist_{ji} ) A_j
$$

Simplified connectivity:

$$
S_i = \sum exp(- dist_{ji} )
$$


## Calculate a simplified connectivity metric for all mapped _Jacaranda_ on BCI
 
First get distances between all trees.
 
```{r}

calculate_dist <- function (data) {

	data %>%
    select(lat, lon) -> plot_matrix

	rdist::pdist(plot_matrix[,c("lat", "lon")], 
	             metric = "euclidean") -> dists
	
	as.data.frame(dists) -> dists_df

	unlist(data$id) -> colnames(dists_df) 

	cbind(data, dists_df)
	
}

distance_df <- calculate_dist(all_jacc)

```

Then calculate the connectivity metric for each tree:

```{r}
all_jacc %>% 
  distinct(id) %>%
  pull(id) -> tree_id_list

calculate_connectivity <- function (data, tree_id) {
  data %>%
    group_by(id) %>%
    mutate(x = exp(- eval(parse(text = tree_id)) ) ) %>%
    ungroup() %>%
    summarise(id = paste(tree_id),
              connectivity = sum(x), .groups = "drop")
}

connectivity_dfs <- lapply(tree_id_list,
                           calculate_connectivity, data = distance_df)

connectivity_dfs %>%
  lapply(drop_na, connectivity) %>%
  dplyr::bind_rows() -> all_connectivity_dfs

```

## Plot

To see if we've covered a good range of the natural densities I'm just going to plot the connectivity values as a histogram.

```{r histogram}
all_connectivity_dfs %>%
  mutate(focal = grepl("EJJACC", id)) -> plot_dat

plot_dat %>%
  ggplot() +
  geom_histogram(aes(x = connectivity, fill = focal))

```

Looks like we could do with some more trees at the lowest levels of connectivity and perhaps some at the very highest. Where can we find those individuals? Lets plot on a map.

```{r map}

bbox <- make_bbox(c(min(all_jacc$lon) - 0.001, max(all_jacc$lon) + 0.005), 
                  c(min(all_jacc$lat) - 0.005, max(all_jacc$lat) + 0.001))

bci_basemap <- ggmap::get_map(bbox, source = "stamen", 
                       force = TRUE, maptype = "toner-lite")

plot_data <- left_join(all_jacc, plot_dat, by = "id")

bci_basemap %>% 
  ggmap() +
  geom_point(data = plot_data, 
             aes(x = lon, y = lat, size = connectivity, colour = focal),
             alpha = 0.5, shape =  16) +
  theme_void()


```

We need to look on the eastern side of the island for low connectivity trees and perhaps pick up a few in/around the 50 ha plot for high connectivity individuals.