---
title: "Make a map of focal trees"
author: "Eleanor Jackson"
date: '`r format(Sys.time(), "%d %B, %Y")`'
always_allow_html: true
output: 
  github_document:
    keep_html: true
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.path = "figures/2022-10-09_create-map/")
```

```{r packages, message=FALSE, warning=FALSE}
library("tidyverse")
library("sf")
library("ggmap")
library("osmdata")
library("plotKML")
```

## Read in gpx files

```{r read-gpx}
plotKML::readGPX(here::here("data", "maps", "focal_jacc.gpx")) %>%
  map_df( ~ .) %>%
  rename(id = name) %>%
  arrange(id) -> focal_jacc

plotKML::readGPX(here::here("data", "maps",
                            "jacc-map-garzonlopez2012", "jac1co_map.gpx")) %>%
  map_df( ~ .) %>%
  select(lon, lat) -> carol_jacc

```

## Make a Jacaranda map

```{r jacc-map, message=FALSE}
bbox <- make_bbox(c(min(focal_jacc$lon) - 0.001, max(focal_jacc$lon) + 0.005), 
                  c(min(focal_jacc$lat) - 0.005, max(focal_jacc$lat) + 0.001))

bci_basemap <- ggmap::get_map(bbox, source = "stamen", 
                       force = TRUE, maptype = "toner-background")
bci_basemap %>% 
  ggmap() +
  geom_point(data = carol_jacc,
             colour = "darkgray",
             size = 1,
           aes(lon, lat)) +
  geom_point(data = focal_jacc,
             colour = "black",
             shape = 4,
             size = 4,
           aes(lon, lat)) +
  theme_void()

```

## Add trails

```{r jacc-map-trails}
# get trails from open street maps
osmdata::opq(bbox = bbox) %>%
  add_osm_feature(key = "highway") %>%
  osmdata_sf() -> osm_trails


bci_basemap %>%
  ggmap() +
  geom_point(data = carol_jacc,
             aes(lon, lat),
             colour = "darkgray",
             size = 1) +
  geom_sf(data = osm_trails$osm_lines,
          inherit.aes = FALSE,
          colour = "forestgreen") +
  geom_point(
    data = focal_jacc,
    aes(lon, lat),
    colour = "black",
    shape = 4,
    size = 4
  ) +
  theme_void()

```

## plot as a density map

```{r density-map}
bci_basemap %>% 
  ggmap() +
  stat_density2d(data = carol_jacc,
                 aes(lon, lat, fill = ..level..), 
                 alpha = 0.5, geom = "polygon") + 
  geom_sf(data = osm_trails$osm_lines,
          inherit.aes = FALSE, colour = "lightgrey") +
  geom_point(data = focal_jacc,
             colour = "red",
           aes(lon, lat)) +
  theme_void()

```

## A map with labels

```{r id-map}
focal_jacc %>%
  mutate(id = gsub("EJJACC", "", focal_jacc$id)) -> focal_jacc_labs

ggplot() +
  geom_sf(data = osm_trails$osm_lines,
          inherit.aes = FALSE,
          colour = "darkgrey") +
  geom_point(data = focal_jacc_labs,
             colour = "red", size = 0.5, 
           aes(lon, lat)) +
  ggrepel::geom_text_repel(data = focal_jacc_labs, 
                           aes(lon, lat, label = id), box.padding = 0.1,
                           colour = "black", hjust = -0.1, size = 2.25) +
  theme_void()

```

## Take a look at canopy gap map

```{r canopy-gaps, message=FALSE}

gaps <- sf::st_read(here::here("data", "maps",
                               "light-gaps-cushman2022", "gaps18to20sp.shp")) 
ggplot() +
  geom_sf(data = gaps, color = "black") +
  theme_void()

```

There are a *lot* of light gaps - let's try filtering them by `area` so we get just the biggest ones. Maybe useful if we can overlay with trails so we can see which trails have gaps on them.

It seems that the canopy gap map doesn't have a coordinate reference system so I can't overlay them on the base map or with the *Jacaranda* density.

```{r canopy-gaps-trails}
ggplot() +
  geom_sf(data = filter(gaps, area > 400), size = 1, color = "black") +
  geom_sf(data = osm_trails$osm_lines,
          inherit.aes = FALSE, colour = "forestgreen") +
  theme_void() +
  ggtitle("Canopy gaps 2018 - 2020")

```


## Exclusion experiment sites

```{r exclusion-sites}

plotKML::readGPX(here::here("data", "maps",
                            "Exclusion_sites_JACC.gpx")) %>%
  map_df( ~ .) %>%
  rename(id = name) %>%
  arrange(id) -> exclusion_sites

bci_basemap %>% 
  ggmap() +
  stat_density2d(data = carol_jacc,
                 aes(lon, lat, fill = ..level..), 
                 alpha = 0.5, geom = "polygon", contour_var = "count", bins = 15) + 
  geom_point(data = exclusion_sites,
             colour = "red",
           aes(lon, lat)) +
  scale_fill_viridis_c() +
  theme_void()

```

