---
title: "Explore capsule sizes"
author: '`r Sys.getenv("USER")`'
date: '`r format(Sys.time(), "%d %B, %Y")`' 
always_allow_html: true
output: 
  github_document:
    keep_html: true
---
	
```{r setup, include = FALSE}
file_name <- rstudioapi::getSourceEditorContext()$path

knitr::opts_chunk$set(
  fig.path =
    paste0("figures/", sub("\\.Rmd$", "", basename(file_name)), "/", sep = "")
)

ggplot2::theme_set(ggplot2::theme_classic(base_size = 10))
```

When collecting data in the field we didn't classify capsules as immature or mature,
but instead recorded their size and whether they were open or not.

From this information we should be able to classify pods as immature or mature.
Immature pods do not dehisce in the canopy but might have opened on the forest floor.

```{r packages, message=FALSE, warning=FALSE}
library("tidyverse")
library("here")
```

```{r data}
read.csv(here::here("data", "raw", "jacaranda_pods.csv"),
         header = TRUE, na.strings = c("", "NA", "missing")) %>%
  filter(tree != "JACC_130") %>%
  mutate(fragment = case_when(str_detect(comments, "fragment") ~ TRUE,
                              TRUE ~ FALSE)) %>%  
  filter(fragment == FALSE) %>% 
  mutate(pod_size_mm = coalesce(pod_size_string_mm, pod_size_mm)) %>%
  mutate(pod_size_mm = as.numeric(pod_size_mm)) -> pod_data
```

```{r open-closed-size}
pod_data %>%
  drop_na(pod_size_mm) %>% 
  ggplot(aes(x = pod_size_mm, fill = pod_open_closed)) +
  geom_density(alpha = 0.6) +
  xlab("Capsule length (mm)") +
  theme(legend.title = element_blank()) +
  geom_vline(xintercept = 40, linetype = 2)
```

That's a really nice bimodal distribution.
40 mm looks like a good cut off for immature fruits. 
There's a small number of pods that are open but are comparatively small 
(the first blue hump). These are probably immature capsules that opened on the forest floor.

Let's take a quick look at capsule sizes between the different categories: mature,
immature and predated.

```{r grouped-pod-size}
pod_data %>%
  mutate(pod_category = case_when(
    pod_size_mm < 40 & str_detect(morph, "^symmetrical_locules") ~ "immature",
    pod_size_mm >= 40 & str_detect(morph, "^symmetrical_locules") ~ "mature",
    morph == "asymmetrical_locules" |
      morph == "single_locule" |
      morph == "small_knobbly" ~ "predated",
    TRUE ~ NA
  )) %>% 
  filter(!is.na(pod_category)) -> pod_groups

pod_groups %>% 
  ggplot(aes(x = pod_size_mm, fill = pod_category)) +
  geom_density(alpha = 0.6) +
  xlab("Capsule length (mm)") +
  theme(legend.title = element_blank())

```

The predated capsules also have a bimodal distribution - this'll be the two different
morphotypes that we found: "small-knobbly" ones that looked like they'd been predated
when they were still young, and "asymmetric" ones which were bigger but had one shrunken locule.

Let's pull some numbers out for the manuscript.

```{r mean-size}
pod_groups %>% 
  group_by(pod_category) %>% 
  summarise(median = median(pod_size_mm), mean = mean(pod_size_mm))
```

