---
title: "Choose an alpha value"
author: "Eleanor Jackson"
date: '23 January, 2023'
always_allow_html: true
output: 
  github_document:
    keep_html: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.path = "figures/2023-01-23_choose-alpha-value/")
ggplot2::theme_set(ggplot2::theme_classic(base_size = 10))
```

To determine the buffer radius to use for our connectivity calculation 
we will fit models with varying values of *r* and compare their AIC values. 
_r_ is the average migration distance of our seed predator and *r* = 1 / $\alpha$

__Connectivity__\

$$C_{i} = \sum exp(-\ \alpha \ dist_{ji})$$

```{r packages}
library("tidyverse")
library("geosphere")
library("sf")
```

## Get data

```{r data}
readRDS(here::here("data", "clean", "pod_data.rds")) %>% 
  mutate(crown_area_m2 = pi * crown_radius_m^2) %>% 
  select(tree_id, lon, lat, crown_area_m2) -> focal_jacc

st_read(here::here(
  "data", "maps",
  "jacc-map-garzonlopez2012", "JAC1COpointSept.shp"
)) %>% 
  filter(!st_is_empty(.)) %>%
  select(geometry) -> point_data

st_read(here::here(
  "data", "maps",
  "polygons-garzonlopez2008", "JAC1CO_pol2008.shp"
)) %>% 
  filter(!st_is_empty(.)) %>%
  select(geometry) -> polygon_data
```

Filter Carol's mapped trees to only include trees that have polygon data 
and use the polygons to calculate crown area (m^2)

```{r}
st_join(polygon_data, point_data, 
        join = st_within, largest = TRUE, left = FALSE) %>%  
  mutate(crown_area_m2 = as.numeric(st_area(geometry))) %>% 
  mutate(tree_id = paste0("CGL_", 1:n()),
         geometry = st_centroid(geometry)) %>% 
  sf::st_transform(crs = "+proj=longlat +datum=WGS84", allow_ballpark = FALSE) %>% 
  mutate(lon = sf::st_coordinates(.)[,1],
         lat = sf::st_coordinates(.)[,2]) -> other_jacc
```


We only want to include reproductive individuals. The minimum reproductive size of *J. copaia* is 20 cm dbh [(Wright et al. 2005)](https://doi.org/10.1017/S0266467405002294).

We can use Helene's allometry data to estimate dbh from crown area 
(see [`bci-jacc/code/exploration/2022-10-11_estimate-crown.md`](bci-jacc/code/exploration/2022-10-11_estimate-crown.md)).

DBH = (crown area + 99.42) / 0.38

```{r}
other_jacc %>%
  mutate(dbh_estimate_mm = (crown_area_m2 + 99.42) / 0.38) %>%
  filter(dbh_estimate_mm >= 200) %>% 
  select(- dbh_estimate_mm) -> other_jacc
```

Oh.. there aren't any trees in the data with a dbh less than 20 cm. 
I guess maybe they're too small to pick up in aerial photographs.

## Remove duplicated trees

Trees mapped in our 'on the ground' data could be the same individuals as those in Carol's aerial maps.

```{r}
geosphere::distm(cbind(pull(focal_jacc, lon), pull(focal_jacc, lat)),
  cbind(pull(other_jacc, lon), pull(other_jacc, lat)),
  fun = distGeo
) -> dist_matrix

rownames(dist_matrix) <- focal_jacc$tree_id
colnames(dist_matrix) <- other_jacc$tree_id

# assume closest CGL tree within 20 m of focal tree is the same tree
as.data.frame.table(dist_matrix, responseName = "dist") %>%
  filter(dist <= 20) %>% 
  group_by(Var1) %>%
  slice(which.min(dist)) %>% 
  ungroup() -> duplicate_trees
  
other_jacc %>%
  filter(!tree_id %in% pull(duplicate_trees, Var2)) -> other_jacc_no_dupes

other_jacc_no_dupes %>% 
  st_drop_geometry() %>% 
  bind_rows(focal_jacc) -> all_jacc

# if focal tree was duplicated in CGL data, use polygon for crown area not radius
duplicate_trees %>%
  left_join(other_jacc, by = c("Var2" = "tree_id")) %>% 
  select(Var1, crown_area_m2) %>% 
  rename(tree_id = Var1) %>%
  right_join(y = all_jacc, by = "tree_id", 
             suffix = c("_poly", "_radius")) %>% 
  mutate(crown_area_m2 = ifelse(is.na(crown_area_m2_poly), 
                                crown_area_m2_radius, crown_area_m2_poly)) %>%
  select(- crown_area_m2_poly, -crown_area_m2_radius) -> all_jacc_areas
```

## Calculate pairwise distances between trees

```{r}
calculate_dist <- function (data) {
  
  data %>%
    select(lon, lat) -> plot_matrix
  
  geosphere::distm(plot_matrix, 
               fun = distGeo) -> dists
  
  as.data.frame(dists) -> dists_df
  
  unlist(data$tree_id) -> colnames(dists_df) 
  
  cbind(data, dists_df)
  
}

distance_df <- calculate_dist(all_jacc_areas)
```

## Calculate connectivity using a range of radii

```{r}

focal_jacc %>%
  distinct(tree_id) %>%
  pull(tree_id) -> focal_tree_id_list

tibble(radii = seq(25, 125, 5)) -> radii_list

lapply(radii_list, function(x) 1 / x ) -> alpha_list

merge(focal_tree_id_list, alpha_list) %>%
  rename(id = x, alpha = radii) -> alpha_tree_id

calculate_connectivity <- function (data, id, alpha) {
  data %>%
    filter(tree_id != eval(parse(text = id))) %>%
    select(id, crown_area_m2) %>%
    mutate(x = exp(- alpha * eval(parse(text = id)) ) * crown_area_m2) %>%
    summarise(connectivity = sum(x),
              tree_id = paste(id),
              alpha = eval(parse(text = paste0(alpha))))
}

connectivity_dfs <- pmap(alpha_tree_id, .f = calculate_connectivity, 
                         data = distance_df)

connectivity_dfs %>%
  dplyr::bind_rows() -> all_connectivity_dfs

readRDS(here::here("data", "clean", "pod_data.rds")) %>% 
  drop_na() %>% 
  mutate_at(c("n_total", "n_predated"), round) %>% 
  left_join(all_connectivity_dfs, by = "tree_id") %>% 
  group_by(alpha) %>%
  nest() %>%
  ungroup() %>%
  pull(data) -> connectivity_dfs_alpha

```

## Fit models and look at AIC 

```{r}

fit_glmms <- function (data) {
  lme4::glmer(
    data,
    formula = cbind(n_predated, (n_total - n_predated)) ~ 
      connectivity_sc + (1 | dbh_mm),
    family = "binomial"
  )
  
}

model_data <- map(connectivity_dfs_alpha, ~ data.frame(.) %>%
                    mutate(connectivity_sc = scale(connectivity)) )

models <- lapply(model_data, fit_glmms)

lapply(models, AIC) -> AIC_list

tibble(radius = unlist(radii_list), AIC = unlist(AIC_list)) -> AIC_df

AIC_df %>%
  ggplot(aes(x = radius, y = AIC)) + geom_point()
  

```

A radius of **`r AIC_df %>% slice(which.min(AIC)) %>% pluck(1,1)`** has the lowest AIC.

I just want to check that we're not missing something by not looking at bigger radii. 
We don't know how far our seed predator or pollinators travel and 
a big part of this study was to look for J-C effects at larger scales.

```{r}
tibble(radii = seq(100, 15000, 500)) -> radii_list

lapply(radii_list, function(x) 1 / x ) -> alpha_list

merge(focal_tree_id_list, alpha_list) %>%
  rename(id = x, alpha = radii) -> alpha_tree_id

connectivity_dfs <- pmap(alpha_tree_id, .f = calculate_connectivity, 
                         data = distance_df)

connectivity_dfs %>%
  dplyr::bind_rows() -> all_connectivity_dfs

readRDS(here::here("data", "clean", "pod_data.rds")) %>% 
  drop_na() %>% 
  mutate_at(c("n_total", "n_predated"), round) %>%
  left_join(all_connectivity_dfs, by = "tree_id") %>% 
  group_by(alpha) %>%
  nest() %>%
  ungroup() %>%
  pull(data) -> connectivity_dfs_alpha

model_data <- map(connectivity_dfs_alpha, ~ data.frame(.) %>%
                    mutate(connectivity_sc = scale(connectivity)))

models <- lapply(model_data, fit_glmms)

lapply(models, AIC) -> AIC_list

tibble(radius = unlist(radii_list), AIC = unlist(AIC_list)) %>%
  ggplot(aes(x = radius, y = AIC)) + geom_point()
```

OK nope! Lots of model convergence warnings too.