---
title: "Test for spatial autocorrelation"
author: "Eleanor Jackson"
date: '20 January, 2023'
always_allow_html: true
output: 
  github_document:
    keep_html: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.path = "figures/2023-01-20_test-spatial-autocorrelation/")
```

```{r packages}
library("tidyverse")
library("sf")
library("ncf")
library("spdep")
library("sp")
  
```

```{r data}
readRDS(here::here("data", "clean", "pod_data.rds")) %>%
  sf::st_as_sf(coords = c("lat", "lon"), crs = 4326, remove = FALSE) %>%
  rowwise() %>%
  mutate(predation_rate = n_predated / n_total) %>%
  drop_na() -> pod_data

```

## Correlogram

A correlogram shows the correlation coefficient for the series lagged (in distance) by one delay at a time. e.g. at lag one you’re looking at the correlation between a point and it’s nearest neighbour. At lag two you’re looking at the correlation between a point and it’s second nearest neighbour.

Values close to 1 indicate clustering while values close to -1 indicate dispersion. A random arrangement would give a value that is close to 0.

```{r correlogram}
# using 10 m increments
plot(ncf::correlog(x = pod_data$lat, y = pod_data$lon, z =  pod_data$predation_rate,
              latlon = TRUE, increment = 0.01))

```

Here, values significant at a nominal (two-sided) 5%-level are represented by filled circles and non-significant values by open circles. 

Most of the plot shows values that are close to zero, without significant auto correlation. There are a few points that do show correlation but they seem quite random and there is not many of them... outliers? I think that if there was a pattern of spatial autocorrelation in the data we would see values that tended towards zero as distance increased.

## Moran's I

The expected value of Moran's I under the null hypothesis of no spatial autocorrelation is `-1/(n-1)`, which equals `-0.01123596` for this data. 

Values significantly below `-1/(n-1)` indicate negative spatial autocorrelation and values significantly above `-1/(n-1)` indicate positive spatial autocorrelation. 

```{r morans-i}
# detect nearest neighbours within a 15 km radius (i.e. whole island)
spdep::dnearneigh(pod_data$geometry, 0, 15, 
                  row.names = pod_data$tree_id,  
                  longlat = TRUE) -> nn

# add spatial weights
lw <- spdep::nb2listw(nn, style = "W", zero.policy = TRUE) 

# Moran's test for spatial autocorrelation
spdep::moran.test(pod_data$predation_rate, lw) 

```

There is very little difference between the expected and observed Moran's I, and the test is non significant, meaning we can accept the null hypothesis of no spatial autocorrelation.

## Variogram

A variogram describes the variance of the difference between values at two locations. 

Here, the x axis represents the distance between traps (km) and each point represents a pair of observations. The distance at which the variogram stops increasing and flattens out is the range. Traps which are closer together than this distance are spatially autocorrelated.

```{r variogram}
sp::coordinates(pod_data) = ~lat+lon

plot(
    gstat::variogram(predation_rate ~ 1, data = pod_data, 
        width = 0.001, cutoff = 15000) 
)

```

Looks to me like we don't have any spatial autocorrelation - we immediately reach the 'sill' where the semivariance is flat.